noam_data_exploration_v1

### 2024-08-05 noam_data_exploration_v3.Rmd
in the middle of creating site level csvs to see if they match J but need to run again.

transect_portfolio_columns <- c("stratum", "Portfolio_effect") # names of columns for dataframe
transect_level_portfolio <- 
  data.frame(matrix(nrow = length(unique(wide_transect_level$stratum)), 
                    ncol = length(transect_portfolio_columns))
             )
colnames(transect_level_portfolio) <- transect_portfolio_columns # change column 
transect_level_portfolio$stratum <- levels(wide_transect_level$stratum) # populate

@@

alpha_diversity$Cover <- rowSums(wide_transect_level[ , -which(names(wide_transect_level) %in% meta_cols)])/100
alpha_diversity$stratum <- paste(alpha_diversity$site, "-", alpha_diversity$habitat, "-", alpha_diversity$transect)
lm_synchrony_transect <- codyn::synchrony(df = transect_level,
                                   time.var = "year",
                                   species.var = "taxa",
                                   abundance.var = "mean_cover",
                                   metric = "Loreau",
                                   replicate.var = "stratum") # Loreau and Mazancourt synchrony
colnames(lm_synchrony_transect) = c("stratum", "Loreau_synchrony") # rename cols for merge later

@@

# stability_transect <- community_stability(df = transect_level,
#                                           time.var = "year",
#                                           abundance.var = "mean_cover",
#                                           replicate.var = "stratum") # using codyn measure of stability
# NOTE: codyn community_stability() is just the inverse of CV
diversity_stability <- alpha_diversity %>%
  dplyr::group_by(site, habitat, site_habitat, transect, stratum) %>%
  dplyr::summarise(richness_mean = mean(richness), 
            richness_se = sd(richness)/sqrt(n()),
            richness_cv = CV(richness),
            shannon_mean = mean(shannon), 
            shannon_se = sd(shannon)/sqrt(n()),
            shannon_cv = CV(shannon),
            shannon_stability = 1/CV(shannon), # Stability as defined by codyn and many others (including Pete in 2022 paper)
            Cover_mean = mean(Cover),
            Cover_se = sd(Cover)/sqrt(n()),
            Cover_cv = CV(Cover),
            Cover_stability = 1/CV(Cover)) # as above

diversity_stability <- merge(diversity_stability, lm_synchrony_transect, by = "stratum") # add in loreau and de Mazancourt synchrony
# diversity_stability <- merge(diversity_stability, gross_synchrony_transect, by = "stratum")
diversity_stability <- merge(diversity_stability, transect_level_portfolio, by = "stratum") # add in portfolio effects
pairs(~ Cover_mean + Cover_cv + Cover_stability + 
        shannon_mean + shannon_cv + shannon_stability + 
        Loreau_synchrony + Portfolio_effect, data = diversity_stability)
```

```{r basic diversity stability plots transect level, fig.height=8, fig.width=10}
(shannon_cv_plot = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = shannon_mean, y = Cover_stability, colour = site, shape = habitat), size = 8) + 
   scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
#    stat_smooth(aes(x = shannon_mean, y = shannon_cv),method = "lm", se = T, size = 2) + 
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

(shannon_cv_plot_2 = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = shannon_mean, y = Cover_stability, colour = site, shape = habitat), size = 8, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~site) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

@@

```{r portfolio effects and synchrony transect level,  fig.height = 8, fig.width = 10}
(loreau_stability_plot = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = Loreau_synchrony, y = Cover_stability, colour = site, shape = habitat), size = 8, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~site) +

@@

# (gross_stability_plot = 
#   ggplot(data = diversity_stability) +
#   geom_point(aes(x = Gross_synchrony, y = Cover_stability, colour = site, shape = habitat), size = 8, stroke = 2) + 
#    scale_colour_manual(values = site_colours) +
#    scale_shape_manual(values = habitat_shapes) +
#   facet_wrap(~site) +
#    
#   theme_bw() + 
#   theme(axis.title.x = element_text(size=18), 
#         axis.title.y = element_text(size=18), 
#         panel.background = element_blank(), 
#         panel.grid.major = element_blank(),  #remove major-grid labels
#         panel.grid.minor = element_blank(),  #remove minor-grid labels
#         plot.background = element_blank(),
#         legend.text = element_text(size = rel(1)), # increase legend text
#         legend.title = element_text(size = rel(1)),
#         legend.position = 'bottom', # move legend to bottom
#         legend.box = 'horizontal', # make legend horizontal
#         legend.box.background = element_rect(colour = "black"))  # add box around legend
#          )
# 
(portfolio_stability_plot = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = Portfolio_effect, y = Cover_stability, colour = site, shape = habitat), size = 8, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~site) +

@@

```{r diversity, portfolio effects and synchrony transect level,  fig.height = 8, fig.width = 10}
(loreau_diversity_plot = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = shannon_mean, y = Loreau_synchrony, colour = site, shape = habitat), size = 8, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~site) +

@@

(portfolio_diversity_plot = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = shannon_mean, y = Portfolio_effect, colour = site, shape = habitat), size = 8, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~site) +

@@

```{r model selection}
diversity_mod_universal <- glmmTMB(Cover_stability ~ Loreau_synchrony + shannon_mean*site*habitat, family = Gamma(link = "log"), data = diversity_stability, na.action = na.pass)
diversity_dredge <- dredge(diversity_mod_universal)
model.sel(diversity_dredge)

# top model is as follows (by a decent margin):
diversity_top <- glmmTMB(Cover_stability ~ Loreau_synchrony + shannon_mean*site + site*habitat, family = Gamma(link = "log"), data = diversity_stability, na.action = na.pass)
summary(diversity_top)
Anova(diversity_top)
```

```{r general cover diversity, fig.height = 30, fig.width = 30}
(diversity_cover_plot = 
  ggplot(data = alpha_diversity) +
  geom_point(aes(x = Cover, y = shannon, colour = site, shape = habitat), size = 8, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~year) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         ) # ABORT ABORT ABORT
```

@@

quadrat_portfolio_columns <- c("location", "Portfolio_effect") # names of columns for dataframe
quadrat_level_portfolio <- 
  data.frame(matrix(nrow = length(unique(data_sub$location)), 
                    ncol = length(quadrat_portfolio_columns))
             )
colnames(quadrat_level_portfolio) <- quadrat_portfolio_columns # change column 
quadrat_level_portfolio$location <- unique(data_sub$location) # populate

@@

# sum(is.na(quadrat_level_portfolio$Portfolio_effect)) # 4 quadrats causing trouble

hist(quadrat_level_portfolio$Portfolio_effect) # some crazy outliers due to low richness

alpha_diversity_quad$Cover <- rowSums(data[ , -which(names(data) %in% meta_cols)])/100
alpha_diversity_quad_sub <- alpha_diversity_quad[!alpha_diversity_quad$location %in% problem_children, ]

# alpha_diversity_quad_sub$Cover <- rowSums(data_sub[ , -which(names(data_sub) %in% meta_cols)])/100
alpha_diversity_quad_sub$stratum <- 
  paste(alpha_diversity_quad_sub$site, "-", alpha_diversity_quad_sub$habitat, "-", alpha_diversity_quad_sub$transect)
lm_synchrony_quadrat <- codyn::synchrony(df = long_data_sub,
                                   time.var = "year",
                                   species.var = "taxa",
                                   abundance.var = "percent_cover",
                                   metric = "Loreau",
                                   replicate.var = "location") # Loreau and Mazancourt synchrony
colnames(lm_synchrony_quadrat) = c("location", "Loreau_synchrony") # rename cols for merge later

@@

community_stability_quadrat <- community_stability(df = long_data_sub,
                                   time.var = "year",
                                   abundance.var = "percent_cover",
                                   replicate.var = "location")
colnames(community_stability_quadrat) = c("location", "Cover_stability")

diversity_stability_quadrat <- alpha_diversity_quad_sub %>%
  dplyr::group_by(location, site, habitat, site_habitat, transect, stratum) %>%
  dplyr::summarise(richness_mean = mean(richness), 
            richness_se = sd(richness)/sqrt(n()),
            richness_cv = CV(richness),
            shannon_mean = mean(shannon), 
            shannon_se = sd(shannon)/sqrt(n()),
            shannon_cv = CV(shannon),
            shannon_stability = 1/CV(shannon), # Stability as defined by codyn and many others (including Pete in 2022 paper)
            Cover_mean = mean(Cover),
            Cover_se = sd(Cover)/sqrt(n()),
            Cover_cv = CV(Cover))

diversity_stability_quadrat <- 
  merge(diversity_stability_quadrat, lm_synchrony_quadrat, by = "location") # add in loreau and de Mazancourt synchrony
# diversity_stability_quadrat <- merge(diversity_stability_quadrat, gross_synchrony_quadrat, by = "location")
diversity_stability_quadrat <- merge(diversity_stability_quadrat, quadrat_level_portfolio, by = "location") # add in portfolio effects
diversity_stability_quadrat <- merge(diversity_stability_quadrat, variance_ratio_quadrat, by = "location")
diversity_stability_quadrat <- merge(diversity_stability_quadrat, community_stability_quadrat, by = "location")
pairs(~ Cover_mean + Cover_cv + Cover_stability + 
        shannon_mean + shannon_cv + shannon_stability + 
        Loreau_synchrony +  Portfolio_effect + VR, data = diversity_stability_quadrat)

@@

```{r cover and cv, fig.height = 8, fig.width = 10}
(cover_cv_plot = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = Cover_mean, y = Cover_cv, colour = site, shape = habitat), size = 4) + 
   scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
   facet_wrap(~site) +

@@

(cover_cv_plot = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = Cover_mean, y = shannon_mean, colour = site, shape = habitat), size = 4) + 
   scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
   facet_wrap(~site) +

@@

```{r basic diversity stability plots quadrat level, fig.height=8, fig.width=10}
(shannon_cv_plot = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = shannon_mean, y = Cover_cv, colour = site, shape = habitat), size = 4) + 
   scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
   facet_wrap(~site) +

@@

(shannon_cv_plot_2 = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = shannon_mean, y = Cover_cv, colour = site, shape = habitat), size = 4, stroke = 1) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~habitat) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

@@

```{r portfolio effects and synchrony transect level,  fig.height = 8, fig.width = 10}
(loreau_stability_plot_quad = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = Loreau_synchrony, y = Cover_cv, colour = site, shape = habitat), size = 4, stroke = 1) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~site) +

@@

# (gross_stability_plot_quad = 
#   ggplot(data = diversity_stability_quadrat) +
#   geom_point(aes(x = Gross_synchrony, y = Cover_stability, colour = site, shape = habitat), size = 8, stroke = 2) + 
#    scale_colour_manual(values = site_colours) +
#    scale_shape_manual(values = habitat_shapes) +
#   facet_wrap(~site) +
#    
#   theme_bw() + 
#   theme(axis.title.x = element_text(size=18), 
#         axis.title.y = element_text(size=18), 
#         panel.background = element_blank(), 
#         panel.grid.major = element_blank(),  #remove major-grid labels
#         panel.grid.minor = element_blank(),  #remove minor-grid labels
#         plot.background = element_blank(),
#         legend.text = element_text(size = rel(1)), # increase legend text
#         legend.title = element_text(size = rel(1)),
#         legend.position = 'bottom', # move legend to bottom
#         legend.box = 'horizontal', # make legend horizontal
#         legend.box.background = element_rect(colour = "black"))  # add box around legend
#          )

(portfolio_stability_plot_quad = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = Portfolio_effect, y = Cover_cv, colour = site, shape = habitat), size = 4, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
 # facet_wrap(~site) +

@@

```{r diversity portfolio effects and synchrony quadrat level,  fig.height = 8, fig.width = 10}
(loreau_diversity_plot_quad = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = shannon_mean, y = Loreau_synchrony, colour = site, shape = habitat), size = 4, stroke = 1) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~site) +

@@

(portfolio_diversity_plot_quad = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = shannon_mean, y = Portfolio_effect, colour = site, shape = habitat), size = 4, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~site) +

@@

```{r diversity models quad level}
diversity_mod_universal <- glmmTMB(Cover_cv ~shannon_mean*site + shannon_mean*habitat + site*habitat + (1|transect/location), family = Gamma(link = "log"), data = diversity_stability_quadrat, na.action = na.pass)
diversity_dredge <- dredge(diversity_mod_universal)
model.sel(diversity_dredge) # best model is as specified
summary(diversity_mod_universal) # low variance in random effects
Anova(diversity_mod_universal) # LR Tests are kinda useless
r.squaredGLMM(diversity_mod_universal) # 0.653, 0.655

synchrony_mod_universal <- glmmTMB(Cover_cv ~Loreau_synchrony*site + Loreau_synchrony*habitat + site*habitat + (1|transect/location), family = Gamma(link = "log"), data = diversity_stability_quadrat, na.action = na.pass)
synchrony_dredge <- dredge(synchrony_mod_universal)
model.sel(synchrony_dredge) # best model is as specified
summary(synchrony_mod_universal)
r.squaredGLMM(synchrony_mod_universal) # 0.800, 0.802

>>>>

transect_portfolio_columns <- c("stratum", "portfolio_effect") # names of columns for dataframe
transect_level_portfolio <- 
  data.frame(matrix(nrow = length(unique(wide_transect_level$stratum)), 
                    ncol = length(transect_portfolio_columns))
             )
colnames(transect_level_portfolio) <- transect_portfolio_columns # change column 
transect_level_portfolio$stratum <- levels(wide_transect_level$stratum) # populate

@@

alpha_diversity$cover <- rowSums(wide_transect_level[ , -which(names(wide_transect_level) %in% meta_cols)])/100
alpha_diversity$stratum <- paste(alpha_diversity$site, "-", alpha_diversity$habitat, "-", alpha_diversity$transect)
lm_synchrony_transect <- codyn::synchrony(df = transect_level,
                                   time.var = "year",
                                   species.var = "taxa",
                                   abundance.var = "mean_cover",
                                   metric = "Loreau",
                                   replicate.var = "stratum") # Loreau and Mazancourt synchrony
colnames(lm_synchrony_transect) = c("stratum", "loreau_synchrony") # rename cols for merge later

@@

# stability_transect <- community_stability(df = transect_level,
#                                           time.var = "year",
#                                           abundance.var = "mean_cover",
#                                           replicate.var = "stratum") # using codyn measure of stability
# NOTE: codyn community_stability() is just the inverse of CV
diversity_stability <- alpha_diversity %>%
  dplyr::group_by(site, habitat, site_habitat, transect, stratum) %>%
  dplyr::summarise(richness_mean = mean(richness), 
            richness_se = sd(richness)/sqrt(n()),
            richness_cv = CV(richness),
            shannon_mean = mean(shannon), 
            shannon_se = sd(shannon)/sqrt(n()),
            shannon_cv = CV(shannon),
            shannon_stability = 1/CV(shannon), # Stability as defined by codyn and many others (including Pete in 2022 paper)
            cover_mean = mean(cover),
            cover_se = sd(cover)/sqrt(n()),
            cover_cv = CV(cover),
            cover_stability = 1/CV(cover)) # as above

diversity_stability <- merge(diversity_stability, lm_synchrony_transect, by = "stratum") # add in loreau and de Mazancourt synchrony
# diversity_stability <- merge(diversity_stability, gross_synchrony_transect, by = "stratum")
diversity_stability <- merge(diversity_stability, transect_level_portfolio, by = "stratum") # add in portfolio effects
pairs(~ cover_mean + cover_cv + cover_stability + 
        shannon_mean + shannon_cv + shannon_stability + 
        loreau_synchrony + portfolio_effect, data = diversity_stability)
```

```{r basic diversity stability plots transect level, fig.height=8, fig.width=10}
(shannon_cv_plot = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = shannon_mean, y = cover_stability, colour = site, shape = habitat), size = 8) + 
   scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
#    stat_smooth(aes(x = shannon_mean, y = shannon_cv),method = "lm", se = T, size = 2) + 
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

(shannon_cv_plot_2 = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = shannon_mean, y = cover_stability, colour = site, shape = habitat), size = 8, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~site) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

@@

```{r portfolio effects and synchrony transect level,  fig.height = 8, fig.width = 10}
(loreau_stability_plot = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = loreau_synchrony, y = cover_stability, colour = site, shape = habitat), size = 8, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~site) +

@@

# (gross_stability_plot = 
#   ggplot(data = diversity_stability) +
#   geom_point(aes(x = Gross_synchrony, y = cover_stability, colour = site, shape = habitat), size = 8, stroke = 2) + 
#    scale_colour_manual(values = site_colours) +
#    scale_shape_manual(values = habitat_shapes) +
#   facet_wrap(~site) +
#    
#   theme_bw() + 
#   theme(axis.title.x = element_text(size=18), 
#         axis.title.y = element_text(size=18), 
#         panel.background = element_blank(), 
#         panel.grid.major = element_blank(),  #remove major-grid labels
#         panel.grid.minor = element_blank(),  #remove minor-grid labels
#         plot.background = element_blank(),
#         legend.text = element_text(size = rel(1)), # increase legend text
#         legend.title = element_text(size = rel(1)),
#         legend.position = 'bottom', # move legend to bottom
#         legend.box = 'horizontal', # make legend horizontal
#         legend.box.background = element_rect(colour = "black"))  # add box around legend
#          )
# 
(portfolio_stability_plot = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = portfolio_effect, y = cover_stability, colour = site, shape = habitat), size = 8, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~site) +

@@

```{r diversity, portfolio effects and synchrony transect level,  fig.height = 8, fig.width = 10}
(loreau_diversity_plot = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = shannon_mean, y = loreau_synchrony, colour = site, shape = habitat), size = 8, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~site) +

@@

(portfolio_diversity_plot = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = shannon_mean, y = portfolio_effect, colour = site, shape = habitat), size = 8, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~site) +

@@

```{r model selection}
diversity_mod_universal <- glmmTMB(cover_stability ~ loreau_synchrony + shannon_mean*site*habitat, family = Gamma(link = "log"), data = diversity_stability, na.action = na.pass)
diversity_dredge <- dredge(diversity_mod_universal)
model.sel(diversity_dredge)

# top model is as follows (by a decent margin):
diversity_top <- glmmTMB(cover_stability ~ loreau_synchrony + shannon_mean*site + site*habitat, family = Gamma(link = "log"), data = diversity_stability, na.action = na.pass)
summary(diversity_top)
Anova(diversity_top)
```

```{r general cover diversity, fig.height = 30, fig.width = 30}
(diversity_cover_plot = 
  ggplot(data = alpha_diversity) +
  geom_point(aes(x = cover, y = shannon, colour = site, shape = habitat), size = 8, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~year) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         ) # ABORT ABORT ABORT
```

@@

quadrat_portfolio_columns <- c("location", "portfolio_effect") # names of columns for dataframe
quadrat_level_portfolio <- 
  data.frame(matrix(nrow = length(unique(data_sub$location)), 
                    ncol = length(quadrat_portfolio_columns))
             )
colnames(quadrat_level_portfolio) <- quadrat_portfolio_columns # change column 
quadrat_level_portfolio$location <- unique(data_sub$location) # populate

@@

# sum(is.na(quadrat_level_portfolio$portfolio_effect)) # 4 quadrats causing trouble

hist(quadrat_level_portfolio$portfolio_effect) # some crazy outliers due to low richness

alpha_diversity_quad$cover <- rowSums(data[ , -which(names(data) %in% meta_cols)])/100
alpha_diversity_quad_sub <- alpha_diversity_quad[!alpha_diversity_quad$location %in% problem_children, ]

# alpha_diversity_quad_sub$cover <- rowSums(data_sub[ , -which(names(data_sub) %in% meta_cols)])/100
alpha_diversity_quad_sub$stratum <- 
  paste(alpha_diversity_quad_sub$site, "-", alpha_diversity_quad_sub$habitat, "-", alpha_diversity_quad_sub$transect)
lm_synchrony_quadrat <- codyn::synchrony(df = long_data_sub,
                                   time.var = "year",
                                   species.var = "taxa",
                                   abundance.var = "percent_cover",
                                   metric = "Loreau",
                                   replicate.var = "location") # Loreau and Mazancourt synchrony
colnames(lm_synchrony_quadrat) = c("location", "loreau_synchrony") # rename cols for merge later

@@

community_stability_quadrat <- community_stability(df = long_data_sub,
                                   time.var = "year",
                                   abundance.var = "percent_cover",
                                   replicate.var = "location")
colnames(community_stability_quadrat) = c("location", "cover_stability")

diversity_stability_quadrat <- alpha_diversity_quad_sub %>%
  dplyr::group_by(location, site, habitat, island_side, site_habitat, transect, stratum) %>%
  dplyr::summarise(richness_mean = mean(richness), 
            richness_se = sd(richness)/sqrt(n()),
            richness_cv = CV(richness),
            shannon_mean = mean(shannon), 
            shannon_se = sd(shannon)/sqrt(n()),
            shannon_cv = CV(shannon),
            shannon_stability = 1/CV(shannon), # Stability as defined by codyn and many others (including Pete in 2022 paper)
            cover_mean = mean(cover),
            cover_se = sd(cover)/sqrt(n()),
            cover_cv = CV(cover))

diversity_stability_quadrat <- 
  merge(diversity_stability_quadrat, lm_synchrony_quadrat, by = "location") # add in loreau and de Mazancourt synchrony
# diversity_stability_quadrat <- merge(diversity_stability_quadrat, gross_synchrony_quadrat, by = "location")
diversity_stability_quadrat <- merge(diversity_stability_quadrat, quadrat_level_portfolio, by = "location") # add in portfolio effects
diversity_stability_quadrat <- merge(diversity_stability_quadrat, variance_ratio_quadrat, by = "location")
diversity_stability_quadrat <- merge(diversity_stability_quadrat, community_stability_quadrat, by = "location")
pairs(~ cover_mean + cover_cv + cover_stability + 
        shannon_mean + shannon_cv + shannon_stability + 
        loreau_synchrony +  portfolio_effect + VR, data = diversity_stability_quadrat)

@@

```{r cover and cv, fig.height = 8, fig.width = 10}
(cover_cv_plot = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = cover_mean, y = cover_cv, colour = site, shape = habitat), size = 4) + 
   scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
   facet_wrap(~site) +

@@

(cover_cv_plot = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = cover_mean, y = shannon_mean, colour = site, shape = habitat), size = 4) + 
   scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
   facet_wrap(~site) +

@@

```{r basic diversity stability plots quadrat level, fig.height=8, fig.width=10}
(shannon_cv_plot = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = shannon_mean, y = cover_cv, colour = site, shape = habitat), size = 4) + 
   scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
   facet_wrap(~site) +

@@

(shannon_cv_plot_2 = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = shannon_mean, y = cover_cv, colour = site, shape = habitat), size = 4, stroke = 1) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~habitat) +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

@@

```{r portfolio effects and synchrony transect level,  fig.height = 8, fig.width = 10}
(loreau_stability_plot_quad = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = loreau_synchrony, y = cover_cv, colour = site, shape = habitat), size = 4, stroke = 1) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~site) +

@@

# (gross_stability_plot_quad = 
#   ggplot(data = diversity_stability_quadrat) +
#   geom_point(aes(x = Gross_synchrony, y = cover_stability, colour = site, shape = habitat), size = 8, stroke = 2) + 
#    scale_colour_manual(values = site_colours) +
#    scale_shape_manual(values = habitat_shapes) +
#   facet_wrap(~site) +
#    
#   theme_bw() + 
#   theme(axis.title.x = element_text(size=18), 
#         axis.title.y = element_text(size=18), 
#         panel.background = element_blank(), 
#         panel.grid.major = element_blank(),  #remove major-grid labels
#         panel.grid.minor = element_blank(),  #remove minor-grid labels
#         plot.background = element_blank(),
#         legend.text = element_text(size = rel(1)), # increase legend text
#         legend.title = element_text(size = rel(1)),
#         legend.position = 'bottom', # move legend to bottom
#         legend.box = 'horizontal', # make legend horizontal
#         legend.box.background = element_rect(colour = "black"))  # add box around legend
#          )

(portfolio_stability_plot_quad = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = portfolio_effect, y = cover_cv, colour = site, shape = habitat), size = 4, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
 # facet_wrap(~site) +

@@

```{r diversity portfolio effects and synchrony quadrat level,  fig.height = 8, fig.width = 10}
(loreau_diversity_plot_quad = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = shannon_mean, y = loreau_synchrony, colour = site, shape = habitat), size = 4, stroke = 1) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~site) +

@@

(portfolio_diversity_plot_quad = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = shannon_mean, y = portfolio_effect, colour = site, shape = habitat), size = 4, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~site) +

@@

```{r diversity models quad level}
diversity_mod_universal <- glmmTMB(cover_cv ~shannon_mean*site + shannon_mean*habitat + site*habitat + (1|transect/location), family = Gamma(link = "log"), data = diversity_stability_quadrat, na.action = na.pass)
diversity_dredge <- dredge(diversity_mod_universal)
model.sel(diversity_dredge) # best model is as specified
summary(diversity_mod_universal) # low variance in random effects
Anova(diversity_mod_universal) # LR Tests are kinda useless
r.squaredGLMM(diversity_mod_universal) # 0.653, 0.655

synchrony_mod_universal <- glmmTMB(cover_cv ~loreau_synchrony*site + loreau_synchrony*habitat + site*habitat + (1|transect/location), family = Gamma(link = "log"), data = diversity_stability_quadrat, na.action = na.pass)
synchrony_dredge <- dredge(synchrony_mod_universal)
model.sel(synchrony_dredge) # best model is as specified
summary(synchrony_mod_universal)
r.squaredGLMM(synchrony_mod_universal) # 0.800, 0.802
